---
title: "The 500 Family Study data"
author: Zhengguo Gu
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
The original raw data -- "500 Family Study" data -- is publicly available at http://www.icpsr.umich.edu/icpsrweb/ICPSR/studies/4549?searchSource=revise&q=500+family+study. To obtain the data used in this article, follow the following steps:

# Step 1. Download data
Download the DS3 and DS4 raw data (SPSS format) from the website. Use SPSS software to save the raw data as comma separated values file. 

# Step 2. Process data

## A function for replacing missing values in the raw data with NA
We first write a simple function for dealing with the coding missing value in the raw data. All missing values are replaced with the mean
```{r}
sim_impu <- function(DATA, M_values, repl){
  # M_values: missing values in the raw data. Some missing values are coded as 9, some are coded as 999 etc. 
  # repl = T, missing values are replaced with mean. 
  
  if(missing(repl)){
    repl = F
  }
  
  nrow <- dim(DATA)[1]
  ncol <- dim(DATA)[2]
  DATA <- data.matrix(DATA)
  
  for(i in 1:length(M_values)){
    DATA[which(DATA == M_values[i])] <- NA
  }
  
  if(repl == T){
    
    for(n in 1:nrow){
    for(m in 1:ncol){
      if(is.na(DATA[n, m])){
          DATA[n, m] <- mean(DATA[n, ], na.rm = T)
      }
    }
   }
  }
  

 return(DATA)
}
```

## Load raw data

```{r}
#DS3 = read.csv("D:/TilburgOffice/Dropbox/tilburg office/Research SCA/Project 2 software Simultaneous/newdata/04549-0003-Data.csv", sep = ";" )
#DS4 = read.csv("D:/TilburgOffice/Dropbox/tilburg office/Research SCA/Project 2 software Simultaneous/newdata/04549-0004-Data.csv", sep = ";" )

DS3 = read.csv("D:/Dropbox/Tilburg office/Research SCA/Project 2 software Simultaneous/newdata/04549-0003-Data.csv", sep = ";" ) #PC at home (synced with dropbox)
DS4 = read.csv("D:/Dropbox/Tilburg office/Research SCA/Project 2 software Simultaneous/newdata/04549-0004-Data.csv", sep = ";" ) #PC at home (synced with dropbox)
```

## Process DS3 data
### Questionnaire #74: Please tell us how much agree with each of the following statements about your relationship with your spouse/parter.

Example question: 
74a. My partner and I understand each other perfectly. (strongly disagree to strongly agree). 


The following questions in this questionnaire are to be reverse coded:


+ b
+ e
+ h
+ i
+ l
+ n

```{r}
DS3[, c(352:367)] <- sim_impu(DS3[, c(352:367)], 9)
DS3[, c(353, 356, 359, 360, 363, 365)] <- 6 - DS3[, c(353, 356, 359, 360, 363, 365)] #reverse coding

DS3[, c(352:367)] <- sim_impu(DS3[, c(352:367)], 9, repl = T) #NA replaced with mean
summary(DS3[, c(352:367)])
```
Note: 96 rows in the data matrix above contain NaN vectors, that is, the entire row is NA.

After reverse coding, **the higher a score, the happier/more satisfied a person in his/her relationship**.

### Questionnaire #75: There are various ways that couples deal with serious disagreements. When you have a serious disagreement with your spouse/partner, how often do you...

Example question:
75a. Just keep your opinions to yourself (Never to very often).


The following question is to be reverse coded:

+ c
+ d

```{r}
DS3[, c(368:371)] <- sim_impu(DS3[, c(368:371)], 9)
DS3[, c(370, 371)] <- 6 - DS3[, c(370, 371)] #so the higher the value the calmer.
DS3[, c(368:371)] <- sim_impu(DS3[, c(368:371)], 9, repl = T)
summary(DS3[, c(368:371)])
```
Note: 96 rows in the data matrix above contain NaN vectors, that is, the entire row is NA.

After reverse coding, **the higher a score, the less violent during a disagreement with the spouse/partner**.

### Questionnaire #80: What do you think the chances are that the child participating in the study will..

Example question: 
a. Graduate from high school (very low to very high).

```{r}
DS3[, c(379:386)] <- sim_impu(DS3[, c(379:386)], c(9, -8), repl = T)
summary(DS3[, c(379:386)])
```
Note: 31 rows in the data matrix above contain NaN vectors, that is, the entire row is NA.

No reverse coding needed. **The higher the score, the stronger the parent feels that the child has a bright future/happy life**.

### Questionnaire #81: How often do you and the child in the study participate in the following activities together:

Example question: 
a. Talking about everyday events in your lives.

```{r}
DS3[, c(387:404)] <- sim_impu(DS3[, c(387:404)], c(9, -8), repl = T)
summary(DS3[, c(387:404)])
```
Note: 33 rows in the data matrix above contain NaN vectors, that is, the entire row is NA.

No reverse coding needed. **The higher the score, the more often the parent does activities together with the child**.

### Questionnaire #85: The statements below describe lots of events that routinely occur in families with young children. These events sometimes make life diffcult. Please read each item and circle how much of a 'hassle' you feel this is for you. If you have more than one child, these events can include any or all of your children. 

Example question: a. Continually clearning up messess of toys or food. 

```{r}
DS3[, c(441:460)] <- sim_impu(DS3[, c(441:460)], c(9, -8), repl = T)
summary(DS3[, c(441:460)])
```
Note: 565 rows in the data matrix above contain NaN vectors, that is, the entire row is NA. We should consider remove this questionnaire. 

No reverse coding needed. 

### Questionnaire #87: Now we're going to present a few more statements about parenting. How true do you feel each of the following statements is in your life?

Example question: a. Being a parent is harder than I thought it would be. (Never true to always true)

```{r}
DS3[, c(540:543)] <- sim_impu(DS3[, c(540:543)], 9, repl = T)
summary(DS3[, c(540:543)])
```
Note: 30 rows in the data matrix above contain NaN vectors, that is, the entire row is NA.

No reverse coding needed. **The higher the score, the more negative feeling about parenting**. 

### Questionnaire #88: In general, I am a parent who...

Example question: a. Makes my children feel better when they talk over their worries with me... (Never true to always true)

```{r}
DS3[, c(544:549)] <- sim_impu(DS3[, c(544:549)], 9, repl = T)
summary(DS3[, c(544:549)])
```
Note: 30 rows in the data matrix above contain NaN vectors, that is, the entire row is NA.

No reverse coding needed. **The higher the score, the more the parent cheers the child up**.

### Questionnaire #89: When you and one of your children have had a disagreement, how often have you...

Example question: a. Just kept your opinion to your self (Never to Very often)

```{r}
DS3[, c(550:553)] <- sim_impu(DS3[, c(550:553)], 9)
DS3[, c(550, 551)] <- 5 - DS3[, c(550, 551)] #note, category 0 to 4
DS3[, c(550:553)] <- sim_impu(DS3[, c(550:553)], 9, repl = T)
summary(DS3[, c(550:553)])
```
Note: 33 rows in the data matrix above contain NaN vectors, that is, the entire row is NA.

After reverse coding, **the higher the score, the more often the parent (aggressively) discusses issues with the child**.

### Questionnaire #91: Now we're going to list statements that deal with ways that you may perceive yourself. Indicate how often these statements apply to you. 

Example question: a. I feel good about myself. 

```{r}
DS3[, c(555:568)] <- sim_impu(DS3[, c(555:568)], 9)
DS3[, c(556:562, 565, 566)] <- 4 -  DS3[, c(556:562, 565, 566)]
DS3[, c(555:568)] <- sim_impu(DS3[, c(555:568)], 9, repl = T)
summary(DS3[, c(555:568)])
```
Note: 29 rows in the data matrix above contain NaN vectors, that is, the entire row is NA.

After the reverse coding, **the higher the score, the more confident (and the less anger) the parent feels about him/herself**. 

### Combining all the questionnaires above
```{r}
DS3_NEW <- rowMeans(DS3[, c(352:367)])
DS3_NEW <- cbind(DS3_NEW, rowMeans(DS3[, c(368:371)]))
DS3_NEW <- cbind(DS3_NEW, rowMeans(DS3[, c(379:386)]))
DS3_NEW <- cbind(DS3_NEW, rowMeans(DS3[, c(387:404)]))
DS3_NEW <- cbind(DS3_NEW, rowMeans(DS3[, c(441:460)]))
DS3_NEW <- cbind(DS3_NEW, rowMeans(DS3[, c(540:543)]))
DS3_NEW <- cbind(DS3_NEW, rowMeans(DS3[, c(544:549)]))
DS3_NEW <- cbind(DS3_NEW, rowMeans(DS3[, c(550:553)]))
DS3_NEW <- cbind(DS3_NEW, rowMeans(DS3[, c(555:568)]))
DS3_NEW <- cbind(DS3[, 1:3],DS3_NEW)

colnames(DS3_NEW)[4:12] <- c("Relationship with partners", "Argue with partners", "Childs bright future",
                               "Activities with children", "Family hassles", "Feeling about parenting", 
                               "Communation with children", "Argue with children", "Confidence about oneself")

summary(DS3_NEW)
```
Note that item "family hassles" contains 565 missing values (i.e., 565 persons did not fill in the famlily hassles questionnaire at all), and thus this item will be removed. 

## Process DS4 data
!! Note: For this dataset, the corresponding questionnaires are not included in the datafile folders, nor are they available online. However, based on the labels in the SPSS file and also in the Code book, we are able to tell which items belong together and form a questionnaire. 
